Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-29 doc/manuscript/ESSDmanuscript.Rmd
small updates

**Abstract.**
Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
We begin by reconciling terminology and mapping ForC fields into EFDB.
This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
At the time of writing, ForC contained ~`r n_ind_records_EFDB_variables` independent records that would be relevant to EFDB, `r n_records_sent_to_EFDB` of which have been submitted to date.
Among the data in ForC, there is disproportionate representation of biomass (particularly aboveground) stocks, with far fewer records for dead organic matter and soil carbon, and relatively few or no records for net annual increments or C fluxes into (gains) or out of (losses) the IPCC-defined C pools.
Geographic representation is also quite uneven, with the highest densities of relevant records in temperate forests, and with relatively scant representation of tropical forests in Africa and Asia. 
ForC represents a diversity of stand ages, although records for young stands are primarily limited to C stocks, as opposed to net increments of fluxes.
This distribution of records is generally reflected in the subset of records that have been submitted to EFDB to date.
In the future, forest C estimates in EFDB can be improved through targeted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

@@

In addition to reviewing existing records, we added a total of `r n_records_added` new records to ForC.
These included **##** records from **##** studies [@piponiot_distribution_2022; @refs<!--others?-->] that were not previously included in ForC.
In addition, we created new records for **##** EFDB-relevant estimates presented in the original publication that were not yet present in ForC.

>>>>

**Abstract.**
Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
We begin by reconciling terminology and mapping ForC fields into EFDB.
This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
At the time of writing, ForC contained ~`r n_ind_records_EFDB_variables` independent records that would be relevant to EFDB, `r n_records_sent_to_EFDB` of which have been submitted to date.
Among the data in ForC, there is disproportionate representation of biomass (particularly aboveground) stocks, with far fewer records for dead organic matter and soil carbon, and relatively few or no records for net annual increments or C fluxes into (gains) or out of (losses) the IPCC-defined C pools.
Geographic representation is also quite uneven, with the highest densities of relevant records in temperate forests, and with relatively scant representation of tropical forests in Africa and Asia. 
ForC represents a diversity of stand ages, although records for young stands are primarily limited to C stocks, as opposed to net increments of fluxes.
This distribution of records is generally reflected in the subset of records that have been submitted to EFDB to date.
In the future, forest C estimates in EFDB can be improved through targeted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.
Given that climate change is rapidly impacting the world's forests, timely reporting of recent estimates will be especially critical to accurate forest C accounting.

@@

In addition to reviewing existing records, we added a total of `r n_records_added` new records to ForC.
These included **##** records from **##** studies [@piponiot_distribution_2022; @lutz_largediameter_2021; @refs<!--others?-->] that were not previously included in ForC.
In addition, we created new records for **##** EFDB-relevant estimates presented in the original publication that were not yet present in ForC.

### 2022-04-29 doc/manuscript/ESSDmanuscript.tex
small updates

\textbf{Abstract.} Forests are critical for climate change mitigation
and consitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). At the time of writing, ForC contained
\textasciitilde17204 independent records that would be relevant to EFDB,
1214 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil carbon,
and relatively few or no records for net annual increments or C fluxes
into (gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. ForC represents a
diversity of stand ages, although records for young stands are primarily
limited to C stocks, as opposed to net increments of fluxes. This
distribution of records is generally reflected in the subset of records
that have been submitted to EFDB to date. In the future, forest C
estimates in EFDB can be improved through targeted research to fill
critical gaps, reporting of information required by IPCC, and continued
submission of data from scientific publications to the EFDB.

@@

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest carbon stocks, increments, and annual
fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. ForC
includes data ingested from individual publications and relevant
databases, including the Global Reforestation Opportunity Assessment
(GROA) database \citep[database doi:
10.5281/zenodo.3983644]{cook-patton_mapping_2020}, the global soil
respiration database
\citep[SRDB-V5,][]{bond-lamberty_global_2010, jian_restructured_2021}.
For C currently contains 39856 records from 10590 plots in 1531 distinct
geographical areas, along with records of stand age and disturbance
history. As such, ForC is positioned to improve forest C accounting
through the transfer of data to EFDB. The purpose of this publication is
to document that process and provide recommendations for future
improvements.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C accounting under IPCC guidelines. However, modest changes to
the structure and contents of ForC were needed in order to provide all
information required by EFDB and to improve ForC's capacity to serve as
a repository of valuable information for forest C accounting under IPCC
guidelines. To support export of data to EFDB, and to improve the
overall quality of the ForC database, we added or modified 18 fields
(Appendix A), defined 15 new variables, implemented enhanced quality
control, manually reviewed \textgreater\# records to obtain additional
required information, and added 325 new records.

@@

In addition to reviewing existing records, we added a total of 325 new
records to ForC. These included \textbf{\#\#} records from \textbf{\#\#}
studies \citep{piponiot_distribution_2022, refs} that were not
previously included in ForC. In addition, we created new records for
\textbf{\#\#} EFDB-relevant estimates presented in the original
publication that were not yet present in ForC.

@@

To transfer complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data transfer were that (1)
records had been checked against the original study and determined to be
complete and correct (\texttt{EFDB.ready} field in
\texttt{ForC\_citations} table), (2) the original study presented values
in tables or text, as opposed to the values having been digitized from
graphs or calculated based on related variables
(\texttt{data.location.within.source} field in
\texttt{ForC\_measurements} table), and (3) the records had not
previously been sent to EFDB. Once converted into EFDB format, the
records were reviewed and then sent to the IPCC's Technical Support Unit
for inclusion in EFDB. Complete records needed to be reviewed by the
EFDB editorial board prior to posting in the database -- a process that
lags behind records transfer and had not yet been completed for all
records sent as of April 27, 2022.

@@

As of April 27, 2022, ForC (v4.0) contained 32550 independent records
(39856 total), 17204 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

As of April 27, 2022, we had reviewed or added 2034 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}

>>>>

\textbf{Abstract.} Forests are critical for climate change mitigation
and consitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). At the time of writing, ForC contained
\textasciitilde17204 independent records that would be relevant to EFDB,
1214 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil carbon,
and relatively few or no records for net annual increments or C fluxes
into (gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. ForC represents a
diversity of stand ages, although records for young stands are primarily
limited to C stocks, as opposed to net increments of fluxes. This
distribution of records is generally reflected in the subset of records
that have been submitted to EFDB to date. In the future, forest C
estimates in EFDB can be improved through targeted research to fill
critical gaps, reporting of information required by IPCC, and continued
submission of data from scientific publications to the EFDB. Given that
climate change is rapidly impacting the world's forests, timely
reporting of recent estimates will be especially critical to accurate
forest C accounting.

@@

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest carbon stocks, increments, and annual
fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. ForC
includes data ingested from individual publications and relevant
databases, including the Global Reforestation Opportunity Assessment
(GROA) database \citep[database doi:
10.5281/zenodo.3983644]{cook-patton_mapping_2020}, the global soil
respiration database
\citep[SRDB-V5,][]{bond-lamberty_global_2010, jian_restructured_2021}.
For C currently contains 39860 records from 10590 plots in 1531 distinct
geographical areas, along with records of stand age and disturbance
history. As such, ForC is positioned to improve forest C accounting
through the transfer of data to EFDB. The purpose of this publication is
to document that process and provide recommendations for future
improvements.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C accounting under IPCC guidelines. However, modest changes to
the structure and contents of ForC were needed in order to provide all
information required by EFDB and to improve ForC's capacity to serve as
a repository of valuable information for forest C accounting under IPCC
guidelines. To support export of data to EFDB, and to improve the
overall quality of the ForC database, we added or modified 18 fields
(Appendix A), defined 15 new variables, implemented enhanced quality
control, manually reviewed \textgreater\# records to obtain additional
required information, and added 329 new records.

@@

In addition to reviewing existing records, we added a total of 329 new
records to ForC. These included \textbf{\#\#} records from \textbf{\#\#}
studies
\citep{piponiot_distribution_2022, lutz_largediameter_2021, refs} that
were not previously included in ForC. In addition, we created new
records for \textbf{\#\#} EFDB-relevant estimates presented in the
original publication that were not yet present in ForC.

@@

To transfer complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data transfer were that (1)
records had been checked against the original study and determined to be
complete and correct (\texttt{EFDB.ready} field in
\texttt{ForC\_citations} table), (2) the original study presented values
in tables or text, as opposed to the values having been digitized from
graphs or calculated based on related variables
(\texttt{data.location.within.source} field in
\texttt{ForC\_measurements} table), and (3) the records had not
previously been sent to EFDB. Once converted into EFDB format, the
records were reviewed and then sent to the IPCC's Technical Support Unit
for inclusion in EFDB. Complete records needed to be reviewed by the
EFDB editorial board prior to posting in the database -- a process that
lags behind records transfer and had not yet been completed for all
records sent as of April 29, 2022.

@@

As of April 29, 2022, ForC (v4.0) contained 32609 independent records
(39860 total), 17204 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

As of April 29, 2022, we had reviewed or added 2034 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}
